Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction

### 2025-06-26 paper.qmd
MNIST example - PH

The digit "1" of the MNIST dataset [@lecun1998] consists of $7877$ grayscale images of handwritten "1"s. Each image is $28 \times 28$ pixels which corresponds to $784$ variables. The first $10$ principal components, explaining $83\%$ of the total variation, are used. This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, meaning that some pixels are almost always dark and some almost always light. We expect that the best layout captures this type of structure and does not exhibit distinct clusters.

>>>>

The digit "1" of the MNIST dataset [@lecun1998] consists of $7877$ grayscale images of handwritten "1"s. Each image is $28 \times 28$ pixels which corresponds to $784$ variables. The first $10$ principal components, explaining $83\%$ of the total variation, are used. This data essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make when sketched. We expect that the best layout captures this type of structure and does not exhibit distinct clusters.
